Holomorphic field theories and higher algebra

### 2023-08-19 highervertexops.tex
first pass at outlining strategy for 2cocycle

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}
\newcommand{\laurent}{\Hat \sfA}
\newcommand{\sfb}{\mathsf{b}}
\newcommand{\sfc}{\mathsf{c}}

@@

\section{Overview of definitions and goalds}

@@

>>>>

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}
\newcommand{\laurent}{\Hat \sfA}
\newcommand{\sfb}{\mathsf{b}}
\newcommand{\sfc}{\mathsf{c}}
\DeclareMathOperator{\colim}{colim}

@@

\section{Overview of definitions and goals}

@@

Let $\CC^\infty$ denote the direct sum of countably many copies of $\CC$.
We will fix a basis $\{x_n\}_{n \in \ZZ}$.
This vector space can also be described as the colimit of finite-dimensional spaces under inclusion 
\[
\CC^{2N+1} \cong \bigoplus_{n = -N}^N \CC x_n \hookrightarrow \bigoplus_{n = -N-1}^{N+1} \CC x_n \cong \CC^{2N+3}.
\]
Later we will identify this vector space with Laurent polynomials $\CC[t,t^{-1}]$ by identifying $t^n$ with~$x_n$.

There are then two distinct infinite-dimensional Lie algebras that naturally act on this space:
\[
\gl_\infty \hookrightarrow \End(\CC^\infty).
\]
Here
\[
\gl_\infty = \colim \gl_{2N+1} = \bigoplus_{m,n \in \ZZ} \Hom(\CC x_m, \CC x_n),
\]
and, in concrete terms, it corresponds to those matrices acting on $\CC^\infty$ that have finitely many nonzero entries.
By contrast, 
\[
\End(\CC^\infty) = \Hom \left(\bigoplus_{m \in \ZZ} \CC x_m, \bigoplus_{n \in \ZZ} \CC x_n \right) \cong \prod_{m\in \ZZ} \bigoplus_{n \in \ZZ} \Hom(\CC x_m, \CC x_n),
\]
and, in concrete terms, it corresponds to those matrices where each column has finitely many nonzero entries but a row can have arbitrarily many nonzero entries.

In between these two sits an interesting Lie algebra $\Bar{\gl}(\infty)$ consisting of matrices with finitely many nonzero {\em diagonals}.

This Lie algebra has two features crucial for us:
\begin{enumerate}
\item It contains the shift operators $S_k: x_n \mapsto x_{n+k}$ and hence contains $\CC[t,t^{-1}]$ as a subalgebra (note that multiplication by $t^k$ corresponds to $S_k$). More generally, every loop algebra $\gl_N(\CC[t,t^{-1}])$ embeds into $\Bar{\gl}(\infty)$.
\item It has a nontrivial 2-cocycle
\end{enumerate}
